---
title: "Project Euler Two"
author: "Luke"
date: "2024-03-20"
categories: [R, Project Euler, C++]
---

# Question Two
Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with $1$ and $2$, the first $10$ terms will be:
1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 
By considering the terms in the Fibonacci sequence whose values do not exceed four million, find the sum of the even-valued terms.

As soon as we hear fibonacci, your brain is probably already going to recursion or dynamic programming. Because loops are slow in R, this means I will go to C++. The C++ and R intersection is pretty nice, I'll show how to use RCpp to accomplish some of this. 

```{cpp}
#include <cmath>
#include <Rcpp.h>
#include <vector>
using namespace Rcpp;

// [[Rcpp::export]]
int even_fibonaccis(int n = 4000000) {
    std::vector<int> fibs;
    fibs.push_back(1);
    fibs.push_back(2);
    long long cur_sum = 2;
    while (true) {
        int next_int = fibs[fibs.size() - 1] + fibs[fibs.size() - 2];
        if (next_int > n) break;
        fibs.push_back(next_int);
        if (next_int % 2 == 0) {
            cur_sum += next_int;
            //Rcout << "  Added to sum: " << next_int << ", Current sum: " << cur_sum << "\n";
        }
    }
    return cur_sum;
}

```

I tested this function first in its own .cpp file, but for the ease of showing it in a notebook, I've included it in a cpp chunk.
```{r}
Rcpp::sourceCpp("/Users/lukeepp/project-euler/pe/pe2.cpp")
system.time(print(even_fibonaccis()))

```

Now the question becomes, can we make it faster?
One optimization we can make is using an array instead of a vector in C++, this statically sized data structure is quite a bit faster to loop over.


```{cpp}
#include <cmath>
#include <Rcpp.h>
#include <vector>
using namespace Rcpp;

// [[Rcpp::export]]
int even_fibonaccis_arr(int n = 4000000) {
    std::array<int, 1000000> fibs;
    fibs[0] = 1;
    fibs[1] = 2;
    long long cur_sum = 2;
    for (int i = 2; i < fibs.size(); ++i) {
        fibs[i] = fibs[i - 1] + fibs[i - 2];
        if (fibs[i] > n) break;
         if (fibs[i] % 2 == 0) {
            cur_sum += fibs[i];
        }
    }
    return cur_sum;
}

```


```{r}
Rcpp::sourceCpp("/Users/lukeepp/project-euler/pe/pe2.cpp")
system.time(print(even_fibonaccis_arr()))

```

These show as the same time on render, but actually the second is quite a bit faster.